### 2008-07-28 doc/storage_system/arc1-storage-documentation.tex
modifications based on the first internal review

\chapter{Design Overview}
\label{sec:overview}

The ARC1 storage system is a distributed system for storing replicated \emph{file}s on several file storage nodes and manage them in a global namespace. The files can be grouped into \emph{collection}s, and a collection can contain sub-collections and sub-sub-collections in any depth. There is a dedicated \emph{root collection} to gather all collections to the global namespace. This hierarchy of collections and files can be referenced using \emph{Logical Name}s (\emph{LN}s). The users can use this global namespace  as they were using a local filesystem. Files can be transfered by multiple transfer protocols, and the client side tools hide this from the user. The replicas of the files are stored on different storage nodes. A storage node should have one or more data transfer service running (e.g.~HTTP(S), FTP(S), GridFTP, ByteIO\footnote{OGSA ByteIO Working Group (BYTEIO-WG), https://forge.gridforum.org/projects/byteio-wg/}, etc.), and one of the services of the ARC1 storage system is needed to manage each storage node and to integrate it into the system. And there are plans to provide a way to integrate third-party storage solutions into the namespace of the ARC1 storage system. The main services are shown on Figure~\ref{fig:services}.

\begin{figure}[ht]
\centering{{\scalebox{0.9}{\includegraphics{arc1-storage-services.pdf}}}
\caption{\label{fig:services}The components of the ARC1 storage;
\emph{A-H}: the \textbf{A-Hash}, which is a currently centralized, later distributed database;
\emph{L}: the \textbf{Librarian}, which stores the metadata and hierarchy of collections and files, the location of replicas, and health data of the Shepherd services, using the A-Hash as database;
\emph{B}: the \textbf{Bartender}, which provides a high-level interface for the users and for other services;
\emph{S}: the \textbf{Shepherd}, which provides a simple interface for storing files on storage nodes.
} }
\end{figure}

@@

In a collection each entry has a name, and this entry can be a sub-collection, in which there are files and sub-sub-collections, etc.

@@

Each VO\footnote{Virtual Organization, http://www.globus.org/grid\_software/security/voms.php} should create a VO-wide root collection, and put it in the generic root collection, e.g.~if a VO called \verb#vo1# creates a collection called \verb#vo1# as a sub-collection of the root collection (which has the GUID \verb#0#), then it can be referred as \verb#0/vo1# or just \verb#/vo1#. Then this VO can create some files, and put them in this \verb#/vo1# collection, e.g.~\verb#/vo1/file1#, etc. Or sub-collections, e.g.~\verb#/vo1/col1#, \verb#/vo1/col2/file3#, etc. For this the VO does not need the install any service. These files and collections can be created using a Bartender service.

@@

\section{The Bartenders} % (fold)
\label{sec:the_bartenders}
Clients can access the storage system through a Bartender service. If a client wants to create a collection, upload or download a file, the first step is to connect to a Bartender. The Bartender then asks a Librarian to resolve Logical Names and return metadata, then initiates file transfers on some storage node by asking the Shepherd service of that node, then returns the transfer URL (\emph{TURL}) to the client, which allows the client to directly transfer from/to the storage node. So the data transfer itself is not going through the Bartender, it is performed over a direct link between a storage node and the client.

There could be any number of independent Bartender services in the system which provides high-availability and load-balancing.

@@

We have a file on our local disk we want to upload to a collection called \verb#/ourvo/common/docs#. (See Figure~\ref{fig:uploading}.)
\begin{enumerate}
\item We contact a Bartender to put the file, we give the size and checksum and other metadata, and the transfer protocols we want to use. And of course we give the Logical Name we want to be the name of the file, which in this case will be \verb#/ourvo/common/docs/proposal.pdf#
\item The Bartender ask a Librarian to traverse this LN.
\item If the Librarian can traverse the whole LN then this LN is already exists, but if the LN is still available and the parent exists, the Librarian’s response contains the GUID and metadata of the parent collection.
\item Then the Bartender creates a new file entry within the Librarian with all the information we gave. 
\item The Librarian returns the GUID of this new entry.
\item Then the Bartender add the name \verb#proposal.pdf# and the new GUID to the collection \verb#/ourvo/common/docs# and from now on there will be a valid LN \verb#/ourvo/common/docs/proposal.pdf# which points to a file which has no replica at all. If someone tried to download the file called \verb#/ourvo/common/docs/proposal.pdf# now, would get an error message `try again later'.
\item The Bartender (using the information system) chooses a Shepherd and gets its endpoint reference. Then the Bartender initiates uploading of the file to the Shepherd: the request includes the size and checksum of the file, the GUID, and the protocols we are able to use.
\item The Shepherd creates a transfer URL and a referenceID for this file and registers the GUID of the file in its own database and reports to the Librarian that there is a new replica with state \verb#creating#. The Librarian gets the message from the Shepherd and creates a new entry in the locations list of the given file with the serviceID and the referenceID the Shepherd have just reported. If someone tries to download this file now, still gets a `try again later' error message, because this new replica is still not \verb#alive#.
\item The Shepherd returns the the TURL to the Bartender.
\item The Bartender returns the TURL to us.
\item Then we can upload the file to this TURL.
\item The Shepherd detects that the file is arrived and reports the change of state to \verb#alive# to the  Librarian who alters the state in the given file-entry. At this point the file has only one replica. 
\end{enumerate}
\begin{itemize}
\item The Shepherd periodically checks the Librarian if this is less than the needed replica number, and if it is then it initiates creating a new replica by a Bartender.
\item The Bartender chooses another Shepherd, initiates the transfer then returns the TURL to the  first Shepherd which uploads the file to the new Shepherd.
\item All the Shepherds check periodically whether their files have enough replica, and if any of them find that there is more replica needed, it initiates creating a new. If more than one Shepherd of course could cause that there will be more replicas than needed. If a Shepherd finds out that a file has more replicas than needed it notifies a Bartender about it.
\item The Bartender ask the Librarian about all Shepherds this file has replicas on, flags this file as `removing a replica' which prevents other Bartenders to remove an other replica accidentally, then make a decision of which one is to be removed, then contacts the chosen Shepherd and asks it to remove the replica. The Shepherd then notify the Librarian, and the Librarian removes the replica, and removes the flag `removing a replica' as well.
\item If we cannot upload the file to the given TURL for some reason, we should remove the file entry from the collection, or we should call \verb#addReplica# to get a new TURL without removing and recreating the file.
\end{itemize}

@@

In both cases the Shepherd service uses two separate directories: one for storing all the files (e.g.~\verb!./store!) and one for the file transfers (e.g.~\verb!./transfer!). The store directory always contains all the files the Shepherd manages, the transfer directory is empty at the beginning.

>>>>

\chapter{Design Overview}
\label{cha:overview}

The ARC1 storage system is a distributed system for storing replicated \emph{file}s on several file storage nodes and manage them in a global namespace.  The files can be grouped into \emph{collection}s (a concept very similar to directories in the common file systems), and a collection can contain sub-collections and sub-sub-collections in any depth. There is a dedicated \emph{root collection} to gather all collections to the global namespace. This hierarchy of collections and files can be referenced using \emph{Logical Name}s (\emph{LN}s). The users can use this global namespace  as they were using a local filesystem. Files can be transfered by multiple transfer protocols, and the client side tools hide this from the user. The replicas of the files are stored on different storage nodes. A storage node here is a network-accessible computer having storage space to share, and one or more data transfer service running (e.g.~HTTP(S), FTP(S), GridFTP, ByteIO\footnote{OGSA ByteIO Working Group (BYTEIO-WG), https://forge.gridforum.org/projects/byteio-wg/}, etc.). For each storage node one of the services of the ARC1 storage system is needed to manage it and to integrate it into the system. And there are plans to provide a way to integrate third-party storage solutions into the namespace of the ARC1 storage system. The main services of the storage system are the following (see Figure~\ref{fig:services}):
\begin{itemize}
    \item the \textbf{A-Hash} service, which is a currently centralized, later distributed database;
    \item the \textbf{Librarian} service, which stores the metadata and hierarchy of collections and files, the location of replicas, and health data of the Shepherd services, using the A-Hash as database;
    \item the \textbf{Bartender} service, which provides a high-level interface for the users and for other services;
    \item the \textbf{Shepherd} service, which provides a simple interface for storing files on storage nodes.
\end{itemize}

\begin{figure}[ht]
\centering{{\scalebox{0.9}{\includegraphics{arc1-storage-services.pdf}}}
\caption{\label{fig:services}The components of the ARC1 storage: the \textbf{A-H}ash service, the \textbf{L}ibrarian service, the \textbf{B}artender service and the \textbf{S}hepherd service.}}
\end{figure}

The descriptions in Chapter~\ref{cha:overview} and Chapter~\ref{cha:use_cases} are containing some elements which are not implemented yet in the first prototype. See Chapter~\ref{cha:technical_description_and_implementation_status} for the details about what is implemented and what is missing.

@@

A collection is basically just a list of entries, where each entry has a name unique within this collection, and an entry can be a file or a sub-collection, in which there are files and sub-sub-collections, etc.

@@

One possible usage scenario would be if each VO (Virtual Organization)\footnote{Virtual Organization, http://www.globus.org/grid\_software/security/voms.php} would create a VO-wide root collection, and put it in the generic root collection, e.g.~if a VO called \verb#vo1# creates a collection called \verb#vo1# as a sub-collection of the root collection (which has the GUID \verb#0#), then it can be referred as \verb#0/vo1# or just \verb#/vo1#. Then this VO can create some files, and put them in this \verb#/vo1# collection, e.g.~\verb#/vo1/file1#, etc. Or sub-collections, e.g.~\verb#/vo1/col1#, \verb#/vo1/col2/file3#, etc. For this the VO does not need the install any service. These files and collections can be created using a Bartender service.

@@

\section{The Bartenders} % (fold)
\label{sec:the_bartenders}

The Bartender service provides a high-level interface for the storage system to the clients. You can create and remove collections, create, get and remove files, move files and collections within the namespace using Logical Names. The Bartender communicates with the Librarian and Shepherd services to accomplish the client’s requests. The actual file data does not go through the Bartender; file transfers are directly performed between the storage nodes and the clients. There could be any number of independent Bartender services in the system which provides high-availability and load-balancing.

@@

We have a file on our local disk we want to upload to a collection called \verb#/ourvo/common/docs#. (See Figure~\ref{fig:uploading}.)
\begin{enumerate}
\item We contact a Bartender to put the file, we give the size and checksum and other metadata, and the transfer protocols we want to use. And of course we give the Logical Name we want to be the name of the file, which in this case will be \verb#/ourvo/common/docs/proposal.pdf#
\item The Bartender ask a Librarian to traverse this LN.
\item If the Librarian can traverse the whole LN then this LN is already exists, but if the LN is still available and the parent exists, the Librarian’s response contains the GUID and metadata of the parent collection.
\item Then the Bartender creates a new file entry within the Librarian with all the information we gave. 
\item The Librarian returns the GUID of this new entry.
\item Then the Bartender add the name \verb#proposal.pdf# and the new GUID to the collection \verb#/ourvo/common/docs# and from now on there will be a valid LN \verb#/ourvo/common/docs/proposal.pdf# which points to a file which has no replica at all. If someone tried to download the file called \verb#/ourvo/common/docs/proposal.pdf# now, would get an error message `try again later'.
\item The Bartender (using the information system) chooses a Shepherd and gets its endpoint reference. Then the Bartender initiates uploading of the file to the Shepherd: the request includes the size and checksum of the file, the GUID, and the protocols we are able to use.
\item The Shepherd creates a transfer URL and a referenceID for this file and registers the GUID of the file in its own database and reports to the Librarian that there is a new replica with state \verb#creating#. The Librarian gets the message from the Shepherd and creates a new entry in the locations list of the given file with the serviceID and the referenceID the Shepherd have just reported. If someone tries to download this file now, still gets a `try again later' error message, because this new replica is still not \verb#alive#.
\item The Shepherd returns the the TURL to the Bartender.
\item The Bartender returns the TURL to us.
\item Then we can upload the file to this TURL.
\item The Shepherd detects that the file is arrived and reports the change of state to \verb#alive# to the  Librarian who alters the state in the given file-entry. At this point the file has only one replica. 
\end{enumerate}
\begin{itemize}
\item The Shepherd periodically checks the Librarian if this is less than the needed replica number, and if it is then it initiates creating a new replica by a Bartender.
\item The Bartender chooses another Shepherd, initiates the transfer then returns the TURL to the  first Shepherd which uploads the file to the new Shepherd.
\item All the Shepherds check periodically whether their files have enough replica, and if any of them find that there is more replica needed, it initiates creating a new. If more than one Shepherd of course could cause that there will be more replicas than needed. If a Shepherd finds out that a file has more replicas than needed it notifies a Bartender about it.
\item The Bartender ask the Librarian about all Shepherds this file has replicas on, flags this file as `removing a replica' which prevents other Bartenders to remove an other replica accidentally, then make a decision of which one is to be removed, then contacts the chosen Shepherd and asks it to remove the replica. The Shepherd then notify the Librarian, and the Librarian removes the replica, and removes the flag `removing a replica' as well.
\item If we cannot upload the file to the given TURL for some reason, we could call \verb#addReplica# to get a new TURL without removing and recreating the file.
\end{itemize}

@@

In both cases the Shepherd and the transfer services should have access to the same local filesystem where the Shepherd creates two separate directories: one for storing all the files (e.g.~\verb!./store!) and one for the file transfers (e.g.~\verb!./transfer!). The store directory always contains all the files the Shepherd manages, the transfer directory is empty at the beginning.
